---
title: "Analysis Questions"
author: Dominik Schneider
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   bookdown::html_document2:
    df_print: paged
    fig_height: 6
    fig_width: 8
    output_dir: reports
    toc: TRUE
    toc_float: TRUE
    toc_depth: 4
    code_download: true
---

```{r setup, include = F}
library(tidyverse)
library(vroom)
library(here)
library(knitr)
require(xtable)
require(lubridate)
walk(dir('src/R','.R$', full.names = T),source)

knitr::opts_chunk$set(echo=F, results=F, fig.width = 10, warning=F, message=F)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


```

# Analysis {.tabset .tabset-fade}

## VIS
- confirm greenness difference by genotype? (use 1 timepoint only)

- greenness difference based on time of day of image? (7/23 - 7/26)

- do we detect plant area diurnal cycle using the 3 times of day (7/23-7/26)

- can we quantify color heterogeneity based on histogram skew/kurtosis? (1 timepoint)


```{r}
data = vroom(here::here('output','vis','vis.csv-single-value-traits.csv')) 
gtypes = read_csv('data/genotype_map.csv')

singledata = full_join(data,gtypes, by = c('plantbarcode','roi')) %>% 
  mutate(imagetime = case_when(hour(timestamp) < 6 ~ 'early',
                               hour(timestamp) < 12 ~ 'morning',
                               hour(timestamp) < 18 ~ 'afternoon'),
         imagetime = factor(imagetime, levels = c('early', 'morning', 'afternoon')),
         dte = as.Date(timestamp))
```

```{r gtypecolorpal}
# sort gtypes so WT is at the start. Easier to assign "black"
singledata$gtype = forcats::fct_relevel(singledata$gtype,'WT',after=0)
gtypeLevels = levels(singledata$gtype)

# setup gtype color palette
n_gtypes = length(gtypeLevels)

gtypecolorpal = c('black',RColorBrewer::brewer.pal(9,'Set1')[1:(n_gtypes-1)])
names(gtypecolorpal) <- gtypeLevels #assign names to colors that match the gtype levels so filtered dataset still uses the same colors for each gtype

```

### Greenness
```{r}
gi = singledata %>% 
  select(timestamp,dte, imagetime, gtype, roi, greenness_index, plantbarcode)
```

Here is the timeseries of greenness for every plant. There is a reasonable amount of spread within each genotype and there seems to be large natural variation.

```{r, eval=F}
ggplot(gi)+
  geom_line(aes(timestamp, y=greenness_index, group=interaction(plantbarcode,roi), color = plantbarcode))+facet_wrap(~gtype)+
  scale_x_datetime(limits = c(ymd('2019-07-23', tz='UTC'), ymd('2019-07-26', tz='UTC')))
```



Here is the median greenness value across all reps separated by date and time of day. There are different number of bioreps depending on the date.

The red bars stand out as higher for npq2 suggesting higher greenness index in the early hours. Otherwise the green and blue bars are about even.

```{r, fig.width = 10}
gi %>% 
  group_by(gtype,imagetime, dte) %>% 
  summarise(med = median(greenness_index,na.rm=T)) %>% 
  ggplot()+
  geom_col(aes(x=dte,y=med,fill=imagetime), position = position_dodge2(width=.9, preserve = "single"))+
  scale_fill_discrete()+
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=1))+
  facet_wrap(~gtype)+
  scale_x_date(date_breaks = '1 day')+
  coord_cartesian(xlim = as.Date(c("2019-07-23","2019-07-29")))
```

Here is a closer look of the trend for each of the timeperiods. Median for each genotype +/- std error. My main observation is that the relative greenness across genotypes is similar at all time points of the day.


```{r, fig.height=10}
gi %>% 
  group_by(gtype,imagetime, dte) %>% 
  summarise(med = median(greenness_index,na.rm=T),
            sd = sd(greenness_index,na.rm=T),
            hi = med + sd/sqrt(n()),
            lo = med - sd/sqrt(n())) %>%
  ggplot()+
  geom_ribbon(aes(x=dte, ymin = lo, ymax=hi, group=gtype), fill='grey20', alpha=0.5)+
  geom_line(aes(dte,med, color=gtype))+
  facet_wrap(~imagetime, nrow=3)+
  scale_color_manual(values = gtypecolorpal)+
    coord_cartesian(xlim = as.Date(c("2019-07-23","2019-07-29")))


```

#### Conclusion 1

Not a large difference between genotypes. Some separation between after 7/28 

#### Conclusion 2

There is some evidence that the early snapshot (~4am) yielded a higher greenness index for npq2 but otherwise the difference is subtle for other genotypes.

### Plant Area

```{r}
pa = singledata %>% 
  select(timestamp,dte, imagetime, gtype, roi, plantarea, plantbarcode)
```
PLant area for each plant.There are some large outliers in nhd1 and npq2. I should investigate the mask for these images. 
 
```{r, eval=F}
ggplot(pa)+
  geom_line(aes(timestamp, y=plantarea, group=interaction(plantbarcode,roi), color = plantbarcode))+
  facet_wrap(~gtype)+
  scale_x_datetime(limits = c(ymd('2019-07-26', tz='UTC'), ymd('2019-07-29', tz='UTC')))+
  coord_cartesian(ylim = c(0,2500))
```

We can look at the plant area for each plant for each timestamp to evaluate a diurnal cycle.

```{r}
pa %>% 
  ggplot()+
  geom_point(aes(timestamp,plantarea), size=1)+
  geom_line(aes(timestamp,plantarea,color=plantbarcode))+
  scale_x_datetime(limits = as.POSIXct(c('2019-07-23','2019-07-30')))+
  facet_wrap(~gtype)+
   theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=1))+
  scale_y_continuous(limits=c(0,2500),expand=c(0,0))
```

Only npq2 has a noticable diurnal size cycle.


A cleaner plot with median plantarea by genotype on each day separated by imaging period more clearly shows a bias based on time of day. Only npq2 shows a strong difference below. 

```{r}
pa %>% 
  group_by(gtype, imagetime, dte) %>% 
  summarise(med = median(plantarea,na.rm=T)) %>% 
  ggplot()+
  geom_line(aes(x=dte,y=med, color=imagetime))+
  facet_wrap(~gtype, scales='free')+
  scale_x_date(date_breaks = '1 day', limits = as.Date(c('2019-07-23','2019-07-29')))+
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=1))+
  scale_y_continuous( expand=c(0,0))
  # coord_cartesian(ylim = c(200,1000))
```

#### Conclusion 1

Imaging time of day affects the plant area only subtley and the signal is not large enough to confidently detect vertical leaf movements. *npq2* has a noticeable increase in size in the early hours but otherwise the difference is lost.


## PSII

- confirm steadstate NPQ and YII difference by genotype

- can we quantify NPQ heterogeneity based on histogram skew/kurtosis? (1 timepoint)

```{r}
data = vroom(here::here('output','psII','output_psII_level1.csv')) 
frames = read_csv(here::here('data','pimframes_map.csv')) %>% filter(frameid > 6) %>% distinct(parameter)

fvfm = data %>% filter(parameter == 'FvFm')
IndC = data %>% filter(parameter != 'FvFm') %>% 
  mutate(parameter = factor(parameter, levels = frames$parameter))

```

The post processing showed that the most reps were imaged betwee 7/23 and 7/29 (~14 reps) before being reduced to ~5 reps (exception WT). I think this is when we swapped in the first round of amiRNA plants in part of the hotel.


First look at how NPQ changes over the course of the experiment. The first thing I notice is the first 3 induction periods exhibit the largest changes in NPQ but after ~80 sec NPQ is relatively steady across the entire experiment. This is expressed quite clearly for WT, kea, nhd1 and glk. npq4 has a very small  rise in npq. npq2 show a very rapid rise in NPQ. There is a noticeable up and down in npq for WT, kea, nhd1, glk but only for steady state (pink, blue lines). In fact the rising npq has a small opposite effect - when ss npq bumped up, the rising npq bumped down. *Was there an environment change on the 25th or 27th?* 


```{r}

IndC %>% 
  # filter(parameter == 't320_ALon' | parameter == 't300_ALon' | parameter == 't280_ALon') %>%
  group_by(parameter,jobdate, gtype) %>% 
  summarize(med=median(npq_avg,na.rm=T)) %>% 
  ggplot()+
  geom_line(aes(x=jobdate, y=med, color = parameter))+
  facet_wrap(~gtype)+
  scale_x_date(date_breaks = '2 day')+
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))+
  scale_x_date(limits=as.Date(c("2019-07-23","2019-07-29")))


```

YII increases more steadily throuhg the induction curve. There is a complementary dip in YII on the 25th and 27th relative to NPQ.

```{r}
IndC %>% 
  # filter(parameter == 't320_ALon' | parameter == 't300_ALon' | parameter == 't280_ALon') %>%
  group_by(parameter,jobdate, gtype) %>% 
  summarize(med=median(yii_avg,na.rm=T)) %>% 
  ggplot()+
  geom_line(aes(x=jobdate, y=med, color = parameter))+
  facet_wrap(~gtype)+
  scale_x_date(date_breaks = '2 day')+
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))+
    scale_x_date(limits=as.Date(c("2019-07-23","2019-07-29")))

```

Below we have the genotype median YII for the induction curves each day. The hook at the beginning means we didn't wait long enough for the first measurement after the AL was turned on?

```{r, fig.height=16}

IndC %>% 
  filter(between(jobdate,as.Date("2019-07-23"),as.Date("2019-07-29"))) %>% 
  group_by(gtype,jobdate,parameter) %>% 
  summarise(med = median(yii_avg,na.rm=T),
            sd  = sd(yii_avg,na.rm=T),
            hi = med+sd/sqrt(n()),
            lo = med-sd/sqrt(n())) %>% 
  mutate(tInd = parse_number(as.character(parameter))) %>% 
  ggplot()+
  geom_ribbon(aes(x=tInd, ymin = lo, ymax = hi), fill = 'grey80')+
  geom_line(aes(x=tInd, y = med))+
  facet_grid(jobdate~gtype)+
  labs(title = "median YII across all reps with se")


```

NPQ response to induction curves look as expected with a quick increase and plateau. 
```{r, fig.height=16}

IndC %>% 
  filter(between(jobdate,as.Date("2019-07-23"),as.Date("2019-07-29"))) %>% 
  group_by(gtype,jobdate,parameter) %>% 
  summarise(med = median(npq_avg,na.rm=T),
            sd  = sd(npq_avg,na.rm=T),
            hi = med+sd/sqrt(n()),
            lo = med-sd/sqrt(n())) %>% 
  mutate(tInd = parse_number(as.character(parameter))) %>%
  ungroup() %>% 
  tidyr::complete(nesting(jobdate = seq(min(jobdate),max(jobdate),by=1)), fill = list(gtype = 'WT')) %>% #use fill to prevent NA facet
  ggplot()+
  geom_ribbon(aes(x=tInd, ymin = lo, ymax = hi), fill = 'grey80')+
  geom_line(aes(x=tInd, y = med))+
  facet_grid(jobdate~gtype)+
    labs(title = "median NPQ across all reps with se")



```

```{r, fig.height=24}

IndC %>% 
  group_by(gtype,jobdate,parameter) %>% 
  summarise(med = median(yii_avg,na.rm=T),
            sd  = sd(yii_avg,na.rm=T),
            hi = med+sd/sqrt(n()),
            lo = med-sd/sqrt(n())) %>% 
  mutate(tInd = parse_number(as.character(parameter))) %>%
  ungroup() %>% 
  tidyr::complete(nesting(jobdate = seq(min(jobdate),max(jobdate),by=1)), fill = list(gtype = 'WT')) %>% #use fill to prevent NA facet
  filter(jobdate >= '2019-07-23', jobdate <= '2019-07-29') %>% 
  ggplot()+
  geom_ribbon(aes(x=tInd, ymin = lo, ymax = hi, group=gtype), fill = 'grey80')+
  geom_line(aes(x=tInd, y = med, color=gtype))+
  facet_grid(jobdate~.)+
  scale_color_manual(values = gtypecolorpal)+
  labs(title = "median YII across all reps with se")

```


Overall this next plot shows NPQ trends as expected from these genotypes. low for npq4. quick rise and steady decline for npq2. WT and nhd1 fairly similar.

```{r, fig.height=24}

IndC %>% 
  group_by(gtype,jobdate,parameter) %>% 
  summarise(med = median(npq_avg,na.rm=T),
            sd  = sd(npq_avg,na.rm=T),
            hi = med+sd/sqrt(n()),
            lo = med-sd/sqrt(n())) %>% 
  mutate(tInd = parse_number(as.character(parameter))) %>%
  ungroup() %>% 
  tidyr::complete(nesting(jobdate = seq(min(jobdate),max(jobdate),by=1)), fill = list(gtype = 'WT')) %>% #use fill to prevent NA facet
  filter(jobdate >= '2019-07-23', jobdate <= '2019-07-29') %>% 
  ggplot()+
  geom_ribbon(aes(x=tInd, ymin = lo, ymax = hi, group=gtype), fill = 'grey80')+
  geom_line(aes(x=tInd, y = med, color=gtype))+
  facet_grid(jobdate~.)+
  scale_color_manual(values = gtypecolorpal)+
  labs(title = "median NPQ across all reps with se")

```

### Conclusion 1



